---
title: "SWORK 881 week02"
author: Ashley Weitensteiner
date: "01/28/2022"
output: html_document
---

# load libraries
```{r setup, include=FALSE}

```

```{r}
library(tidyverse)
library(tidymodels)
```


# Data
```{r}
# load data
give_your_data_name_here<- read_csv("")

datacc<-read_csv("Data/LHS_exercise_01272022.csv")

#how many observations and variables? 

```
#how many observations and variables? 
# 268 observations and 8 variables


```{r}
# create a new variable called `depression` by summing all ptsd variables. Use tidy approach

dataccD<-datacc%>%
  mutate(depression= rowSums(across(ptsd1_1:ptsd4_1), na.rm = TRUE))
```


```{r}
#run simple descriptive statistics for all variables
mean(dataccD$age)
sd(dataccD$age)
mean(dataccD$depression)
sd(dataccD$depression)
ftable(dataccD$sex)
ftable(dataccD$education)
```
#Age 
32.11 (12.48)
mean(sd)

#depression 
mean score of 8.97, standard deviation of 4.39

#Sex
0 - female = (n = 164/268 = 61.12%)
1 - male = (n = 104/268 = 38.81%)

#Education
1- less than high school = (n = 18/268 = 06.72%)
2 - graduated high school = (n = 91/268 = 33.96%)
3 - some college = (n = 86/268 = 32.09%)
4 - graduated college = (n = 73/268 = 27.24%)


DESCRIBE THE DATA BRIEFLY HERE
Of the sample of 268 participants, 61.12% were female. The mean age of the particpant was 32 years old. Nearly all particpants (95.28%) graduated high school and 27.24% had completed college. The mean depression score was 8.97.



# Model and Visualization
> What is the effect of `age`, `sex`, and `education` on `depression score`? 

Create a linear regression model that answers the question above. Print the results of the output and model diagnostics.

```{r}
# lm
#age
fit.1 <- lm(depression ~ age+factor(education)+factor(sex), data = dataccD)
summary(fit.1)

```

```{r}
par(mfrow=c(2,2))
plot(fit.1)
```

```{r}
# use ggplot to plot histogram of residuals (use 20 bins)

ggplot(fit.1,
       aes(fit.1$residuals))+
  geom_histogram(bins = 20)
```

# Discussion of Results

Describe and interpret your results, including any statistical information you believe important:
# the histogram shows that the data is heavy-tailed, thus the condition to meet normality is not met. This is also reflected in the normal qq plot, in which the residuals deviate from the line in the bottom tail and some in the top tail. 

#of the dependent variables(sex, education, and age) that we tested, only age demonstrated a significant relationship with depression. Age and depression score were negatively associated.
